---
title: "MFA Vignette"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Load the package:
```{r}
library(mfa)
```

## Construct Function

Here is how to create an MFA object based on the wine dataset:
```{r}
# read data
url="https://raw.githubusercontent.com/ucb-stat243/stat243-fall-2016/master/problem-sets/final-project/data/wines.csv" 
data=read.csv(file=url, header = TRUE, sep = ",")
# exclude the first column since it is not numeric
datanum <- data[,-1]

# a list of numeric vectors with the position of the active variables
# to partition the grand table into 10 subtables
sets <- list(1:6, 7:12, 13:18, 19:23, 24:29, 30:34, 35:38, 39:44, 45:49, 50:53)

# create object 'wine' of class mfa
wine <- mfa_gen(data = datanum, sets = sets, ncomps = 2)
```

The wine object is a list with the following elements:
- "sets": user-defined list of vectors with the position of active variables to partition the grand table into subtables
- "eigenvalues": vector of eigenvalues with the length of user-defined `ncomps`
- "FactorScore": factor score of Generalized Singular Value Decomposition (GSVD)
- "Loadings" : variable loadings of GSVD
- "PartialFactorScores": list of partical factor scores of subtables
- "MatrixA": diagnal matrix of the inversed squared singular values 


## Print Method

The print method is designed to print out an MFA object's eigenvalues, and partial factor scores of a subtable if the user specifies the index of the subtable as the argument `subtale`. By default, `subtable = NULL`, and the partial factore scores are not presented.

Here is how to print the eigenvalues of the MFA object wine:
```{r}
print(wine)
```

Here is how to print the eigenvalues and the partical factor scores for the $6^{th}$ subtable:
```{r}
print(wine, subtable = 6)
```


## Plot Method

Given the indices of extracted PCs as 2 dimensions (argument `dim1`, `dim2`), the plot method plots graphs of:
- factor scores (the compromise of the tables) if argument `type = 1`
- partial factor scores if argument `type = 2`
- variable loadings if argument `type = 3`
- partial factor scores and variable loadings if argument `type = 4`

To adjust the scale of the graphs, change the value of arguments `scale_x`, `scale_y`. 

The `text` argument allows users to define lables for variables as a vector.

The `cat` argument allows users to define categories to group the observations as a vector.

The `var` argument allows users to indicate the graphs of specific subtables as a numerical vector of the indices of subtables.

Here is how to plot the partial factor scores (type = 2) for the 8th subtable with the categories to group the observations:
```{r, dpi=180, fig.align='center'}
library(ggplot2)
library(grid)
library(gridExtra)

# define the categories to group the observations
cat <- list("New Zealand" = 1:4, "France" = 5:8, "Canada" = 9:12)

plot(mfa = wine, dim1 = 1, dim2 = 2, type = 2, cat = cat, var = 8)
```

Here is how to plot the variable loadings (type = 3) for the 3rd subtable without labels for variables:
```{r, dpi=180, fig.align='center'}
plot(mfa = wine, dim1 = 1, dim2 = 2, type = 3, var = 3)
```

Plot the variable loadings (type = 3) for the 6th subtable with labels (defined by users) for variables:
```{r, dpi=180, fig.align='center'}
# define the labels for variables
dftext <- c("Cat Pee", "Passion Fruit", "Green Pepper", "Mineral", 
            "Smoky", "Cirtrus", "Cat Pee", "Passion Fruit", "Green Pepper",
            "Mineral","Tropical", "Leafy", "Cat Pee", "Passion Fruit", 
            "Green Pepper", "Mineral", "Grassy", "Flinty", "Cat Pee", 
            "Passion Fruit", "Green Pepper", "Mineral", "Leafy", "Cat Pee",
            "Passion Fruit", "Green Pepper", "Mineral", "Vegetal", "Hay", 
            "Cat Pee", "Passion Fruit", "Green Pepper", "Mineral", "Melon", 
            "Cat Pee", "Passion Fruit", "Green Pepper", "Mineral", "Cat Pee", 
            "Passion Fruit", "Green Pepper", "Mineral", "Grass", "Smoky", 
            "Cat Pee", "Passion Fruit", "Green Pepper", "Mineral", "Peach", 
            "Cat Pee", "Passion Fruit", "Green Pepper", "Mineral")

plot(mfa = wine, dim1 = 1, dim2 = 2, type = 3, text = dftext, var = 6)
```

Plot the partial factor scores and variable loadings (type = 4) for the 1st subtable with the categories to group the observations:
```{r, dpi=180, fig.align='center'}
plot(mfa = wine, type = 4, cat = cat, scale_x = 1/0.8, scale_y = 1/0.4, var = 1)
```


## Eigenvalue Summary

The `eigenval` method creates a table that summarizes information about the obtained eigenvalues of the object wine. 

Here is how to use it:
```{r}
eigenval(wine)
```


## Contributions

The `contri_obs` method calculates and present the contribution of observations to the extracted components (dimensions). The [i,j] element of the printed matrix is the contribution of the ith observation to the jth component. 

Here is how to use it:
```{r}
contri_obs(wine)
```

The `contri_var` method calculates and present the contribution of variables to the extracted components (dimensions). The [i,j] element of the printed matrix is the contribution of the ith variable to the jth component. 

Here is how to use it:
```{r}
contri_var(wine)
```

The `contri_table` method calculates and present the contribution of subtables to the extracted components (dimensions). The [i,j] element of the printed matrix is the contribution of the ith subtable to the jth component. 

Here is how to use it:
```{r}
contri_table(wine)
```


## RV coefficients and Lg coefficients

RV() function calculates the Rv coefficient between 2 specific tables.

```{r}
# define 2 tables
t1 <- matrix(c(2, 4, 3, 1, 5, 7))
t2 <- matrix(c(1, 8, 9, 10, 4, 6))

# calculate the Rv coefficient between t1 and t2
RV(table1 = t1, table2 = t2)
```


To calculate the Rv coefficients between the specified subtables in a grand table (datanum):
```{r}
RV_table(datanum, sets = list(sets[[1]], sets[[2]]))
```


To calculate the Lg coefficients between the specified subtables in a grand table (datanum):
```{r}
Lg_table(datanum, sets = list(sets[[1]], sets[[2]]))
```

## Bootstrap

The bootstrap function performs bootstrap by `L` times, and get L estimated bootstrapped factor scores. The grand table `dataset` is partitioned by argument `userset`. The bootstrap ratios of observations are calculated accordingly. These ratios can be used to find the observation that reliably contribute to a given component. 

Here is how to perform 1000 bootstrap and get the bootstrap ratios of the 12 wines.
```{r}
# a list of numeric vectors with the position of the active variables
# to partition the grand table into 10 subtables
sets <- list(1:6, 7:12, 13:18, 19:23, 24:29, 30:34, 35:38, 39:44, 45:49, 50:53)

bootstrap(L = 1000, userset = sets, dataset = datanum, ncomp = 2)
```
